---
title       : Insert the chapter title here
description : Insert the chapter description here

--- type:ShinyNormalExercise lang:r xp:100 skills:1
## My first app

Assignment comes here. Use Markdown for text formatting.

*** =instructions
- instruction 1
- instruction 2

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# Pre exercise code possible if you want to do some work that the student doesn't see.
# Like pre-loading datasets to variables.
```

*** =sample_code
```{r ui.R}
library(shiny)

# Define UI for application that draws a histogram
ui <- shinyUI(fluidPage(

   # Application title
   titlePanel("Old Faithful Geyser Data"),

   # Sidebar with a slider input for number of bins
   sidebarLayout(
      sidebarPanel(
         ___
      ),

      # Show a plot of the generated distribution
      mainPanel(
         ___
      )
   )
))
```

```{r server.R}
server <- shinyServer(function(input, output) {

   output$distPlot <- renderPlot({
      # generate bins based on input$bins from ui.R
      x    <- faithful[, 2]
      bins <- ___

      # draw the histogram with the specified number of bins
      hist(x, breaks = bins, col = 'darkgray', border = 'white')
   })
})
```

*** =solution
```{r ui.R}
library(shiny)

# Define UI for application that draws a histogram
ui <- shinyUI(fluidPage(

   # Application title
   titlePanel("Old Faithful Geyser Data"),

   # Sidebar with a slider input for number of bins
   sidebarLayout(
      sidebarPanel(
         sliderInput("bins",
                     "Number of bins:",
                     min = 1,
                     max = 50,
                     value = 30)
      ),

      # Show a plot of the generated distribution
      mainPanel(
         plotOutput("distPlot")
      )
   )
))
```

```{r server.R}
server <- shinyServer(function(input, output) {

   output$distPlot <- renderPlot({
      # generate bins based on input$bins from ui.R
      x    <- faithful[, 2]
      bins <- seq(min(x), max(x), length.out = input$bins + 1)

      # draw the histogram with the specified number of bins
      hist(x, breaks = bins, col = 'darkgray', border = 'white')
   })
})
```

*** =sct
```{r}
success_msg("Great work!")
```

--- type:ShinyNormalExercise lang:r xp:100 skills:1
## My first app (2)

Assignment comes here. Use Markdown for text formatting.

*** =instructions
- instruction 1
- instruction 2

*** =hint
hint comes here

*** =pre_exercise_code
```{r}
# Pre exercise code possible if you want to do some work that the student doesn't see.
# Like pre-loading datasets to variables.
```

*** =sample_code
```{r app.R}
library(shiny)

# Define UI for application that draws a histogram
ui <- shinyUI(fluidPage(

   # Application title
   titlePanel("Old Faithful Geyser Data"),

   # Sidebar with a slider input for number of bins
   sidebarLayout(
      sidebarPanel(
         ___
      ),

      # Show a plot of the generated distribution
      mainPanel(
         ___
      )
   )
))

server <- shinyServer(function(input, output) {

   output$distPlot <- renderPlot({
      # generate bins based on input$bins from ui.R
      x    <- faithful[, 2]
      bins <- ___

      # draw the histogram with the specified number of bins
      hist(x, breaks = bins, col = 'darkgray', border = 'white')
   })
})

shinyApp(ui = ui, server = server)
```

*** =solution
```{r app.R}
library(shiny)

# Define UI for application that draws a histogram
ui <- shinyUI(fluidPage(

   # Application title
   titlePanel("Old Faithful Geyser Data"),

   # Sidebar with a slider input for number of bins
   sidebarLayout(
      sidebarPanel(
         sliderInput("bins",
                     "Number of bins:",
                     min = 1,
                     max = 50,
                     value = 30)
      ),

      # Show a plot of the generated distribution
      mainPanel(
         plotOutput("distPlot")
      )
   )
))

server <- shinyServer(function(input, output) {

   output$distPlot <- renderPlot({
      # generate bins based on input$bins from ui.R
      x    <- faithful[, 2]
      bins <- seq(min(x), max(x), length.out = input$bins + 1)

      # draw the histogram with the specified number of bins
      hist(x, breaks = bins, col = 'darkgray', border = 'white')
   })
})

shinyApp(ui = ui, server = server)
```

*** =sct
```{r}
success_msg("Great work!")
```
