---
title: Customizing appearance
description: ''

--- type:VideoExercise lang:r xp:50 skills:1 key:6ea0da9f22
## Interface builder functions

Demo a few of the simpler interface builder options with HTML code

*** =video_link
//player.vimeo.com/video/154783078

*** =video_hls
//videos.datacamp.com/transcoded/000_placeholders/v1/hls-temp.master.m3u8

--- type:NormalExercise lang:r xp:100 skills:1 key:74f15f391f
## Add text with HTML tags in the main panel

Add explanatory tags with HTML tags in the main panel

*** =instructions

*** =hint

*** =pre_exercise_code
```{r}
library(shiny)
```

*** =sample_code
```{r}

```

*** =solution
```{r}

```

*** =sct
```{r}

```

--- type:NormalExercise lang:r xp:100 skills:1 key:98de7f6453
## Add image with img tag

THIS WON'T WORK WITHOUT WWW FOLDER, IS THAT COMING?

*** =instructions
- Test
*** =hint

*** =pre_exercise_code
```{r}
library(shiny)
```

*** =sample_code
```{r}

```

*** =solution

```{r}
library(shiny)

ui <- fluidPage(
  titlePanel("Add 2"),
  tags$img(height = 100, width = 300, src = "http://s3.amazonaws.com/assets.datacamp.com/production/course_4850/datasets/roles_implement.png"),
  sidebarLayout(
    sidebarPanel( sliderInput("x", "Select x", min = 1, max = 50, value = 30) ),
    mainPanel( textOutput("x_updated") )
  )
)

server <- function(input, output, session) {
  add_2            <- function(x) { x + 2 }
  current_x        <- reactive({ add_2(input$x) })
  output$x_updated <- renderText({ current_x() })
}

shinyApp(ui, server)
```

*** =sct
```{r}

```

--- type:VideoExercise lang:r xp:50 skills:1 key:19d1823e6b
## Layout panels

sidebarPanel(), mainPanel(), titlePanel(), conditionalPanel()

*** =video_link
//player.vimeo.com/video/154783078

*** =video_hls
//videos.datacamp.com/transcoded/000_placeholders/v1/hls-temp.master.m3u8

--- type:NormalExercise lang:r xp:100 skills:1 key:4b74aa1cb1
## Extend the sidebar

Add input elements to the sidebar

*** =instructions

*** =hint

*** =pre_exercise_code
```{r}
library(shiny)
```

*** =sample_code
```{r}

```

*** =solution
```{r}

```

*** =sct
```{r}

```

--- type:NormalExercise lang:r xp:100 skills:1 key:655064cc38
## Extend the mainPanel

Add input elements to the main panel

*** =instructions

*** =hint

*** =pre_exercise_code
```{r}
library(shiny)
```

*** =sample_code
```{r}

```

*** =solution
```{r}

```

*** =sct
```{r}

```

--- type:NormalExercise lang:r xp:100 skills:1 key:6dd3650aea
## Add a conditional panel

Add a conditional panel based on user input selection

*** =instructions

*** =hint

*** =pre_exercise_code
```{r}
library(shiny)
```

*** =sample_code
```{r}

```

*** =solution
```{r}

```

*** =sct
```{r}

```

--- type:VideoExercise lang:r xp:50 skills:1 key:45fac2beba
## Tabs and tabset panels

Distributing contents across tabs with tabPanel() and tabsetPanel()

*** =video_link
//player.vimeo.com/video/154783078

*** =video_hls
//videos.datacamp.com/transcoded/000_placeholders/v1/hls-temp.master.m3u8

--- type:NormalExercise lang:r xp:100 skills:1 key:54691fb979
## Create new tab for displaying data table

Create new tab for displaying data table

*** =instructions

*** =hint

*** =pre_exercise_code
```{r}
library(shiny)
```

*** =sample_code
```{r}

```

*** =solution
```{r}

```

*** =sct
```{r}

```

--- type:NormalExercise lang:r xp:100 skills:1 key:77f5943403
## Create new conditional tab for displaying data table

Display the data table in new tab only if checkbox is checked

*** =instructions

*** =hint

*** =pre_exercise_code
```{r}
library(shiny)
```

*** =sample_code
```{r}

```

*** =solution
```{r}

```

*** =sct
```{r}

```

--- type:NormalExercise lang:r xp:100 skills:1 key:8e7129f5e1
## Add new plot to new tab

More extensive extension with new plot in a new tab

*** =instructions

*** =hint

*** =pre_exercise_code
```{r}
library(shiny)
```

*** =sample_code
```{r}

```

*** =solution
```{r}

```

*** =sct
```{r}

```




--- type:NormalExercise lang:r xp:100 skills:1 key:c16e760e4d
## Customize the appearance of your app with shinythemes

Pick a theme and apply to existing app

https://rstudio.github.io/shinythemes/

*** =instructions

*** =hint

*** =pre_exercise_code
```{r}
library(shiny)
```

*** =sample_code
```{r}

```

*** =solution
```{r}

```

*** =sct
```{r}

```


--- type:VideoExercise lang:r xp:50 skills:1 key:73736608ba
## Congratulations, your first app is complete!

First app is complete and ready for sharing, and list some reactivity best practices, Demo local app running, shinyapps.io, Shiny Server Pro, and RStudio Connect options for deploying apps

*** =video_link
//player.vimeo.com/video/154783078

*** =video_hls
//videos.datacamp.com/transcoded/000_placeholders/v1/hls-temp.master.m3u8
