---
title: Introduction and Shiny basics
description: Architecture of a shiny app

--- type:VideoExercise lang:r xp:50 skills:1 key:53592995dc
## Welcome to the course!

01-01-welcome

*** =video_link
//player.vimeo.com/video/154783078

*** =video_hls
//videos.datacamp.com/transcoded/000_placeholders/v1/hls-temp.master.m3u8

--- type:MultipleChoiceExercise lang:r xp:50 skills:1 key:0d65e2ca2d
## What's in an app?

Which is not generally a part of the Shiny app architecture?

*** =instructions
- A function that installs an R package
- User interface
- Server function
- A function that creates Shiny app objects

*** =hint


*** =pre_exercise_code
```{r}
library(shiny)
```

*** =sct
```{r}
test_mc(1)
```

--- type:NormalExercise lang:r xp:100 skills:1 key:725a80bb52
## First peek under the hood

On the right you can see the complete code to reproduce the app we demoed in the previous video. Now you get to interact with the app yourself, and make small adjustments to it. 

*** =instructions

- Hit Submit Answer to run the code and generate the app.
- Play with the input sliders and observe how the output changes.
- Change the `selected` choice for the `y` input to `imdb_rating` and for the `x` input to `imdb_num_votes` and rerun the app.

*** =hint

*** =pre_exercise_code
```{r}
library(shiny)
library(ggplot2)
```

*** =sample_code
```{r}
library(shiny)
library(ggplot2)
load(url("http://s3.amazonaws.com/assets.datacamp.com/production/course_4850/datasets/movies.Rdata"))

# Define UI for application that plots features of movies 
ui <- fluidPage(
  
  # Sidebar layout with a input and output definitions 
  sidebarLayout(
    
    # Inputs: Select variables to plot
    sidebarPanel(
      
      # Select variable for y-axis
      selectInput(inputId = "y", 
                  label = "Y-axis:",
                  choices = c("imdb_rating", "imdb_num_votes", "critics_score", "audience_score", "runtime"), 
                  selected = "audience_score"),
      # Select variable for x-axis
      selectInput(inputId = "x", 
                  label = "X-axis:",
                  choices = c("imdb_rating", "imdb_num_votes", "critics_score", "audience_score", "runtime"), 
                  selected = "critics_score")
    ),
    
    # Output: Show scatterplot
    mainPanel(
      plotOutput(outputId = "scatterplot")
    )
  )
)

# Define server function required to create the scatterplot
server <- function(input, output, session) {

  # Create scatterplot object the plotOutput function is expecting
  output$scatterplot <- renderPlot({
    ggplot(data = movies, aes_string(x = input$x, y = input$y)) +
      geom_point()
  })
}

# Create a Shiny app object
shinyApp(ui = ui, server = server)
```

*** =solution
```{r}

```

*** =sct
```{r}

```

--- type:VideoExercise lang:r xp:50 skills:1 key:fbe44280e8
## User interface

01-02-ui

*** =video_link
//player.vimeo.com/video/154783078

*** =video_hls
//videos.datacamp.com/transcoded/000_placeholders/v1/hls-temp.master.m3u8

--- type:NormalExercise lang:r xp:100 skills:1 key:cf62e7bcf9
## Extend the UI

We'll start our learning with a simplified version of the app you saw in the previous exercise. In this app a `selectInput` widget is used to allow the user to select which variables should be plotted on the `x` and `y` axes of the scatterplot.

The `selectInput` function has four arguments: an `inputId` that is used to refer to the input parameter when building the scatterplot, a list of `choices` to pick from, and a `selected` choice for when the app first launches. Note that `choices` takes a named vector, and the name rather than the value (which must match variable names in the data frame) is displayed to the user. 

*** =instructions

- Add a new `selectInput` widget to color the points by a choice of the following variables: `"title_type"`, `"genre"`, `"mpaa_rating"`, `"critics_rating"`, `"audience_rating"`.
- Make the default selection `"mpaa_rating"`.

*** =hint

*** =pre_exercise_code
```{r}
library(shiny)
library(ggplot2)
```

*** =sample_code
```{r}
library(shiny)
library(ggplot2)
load(url("http://s3.amazonaws.com/assets.datacamp.com/production/course_4850/datasets/movies.Rdata"))

# Define UI for application that plots features of movies
ui <- fluidPage(
  
  # Sidebar layout with a input and output definitions
  sidebarLayout(
    
    # Inputs: Select variables to plot
    sidebarPanel(
      
      # Select variable for y-axis
      selectInput(inputId = "y", 
                  label = "Y-axis:",
                  choices = c("imdb_rating", "imdb_num_votes", "critics_score", "audience_score", "runtime"), 
                  selected = "audience_score"),
      
      # Select variable for x-axis
      selectInput(inputId = "x", 
                  label = "X-axis:",
                  choices = c("imdb_rating", "imdb_num_votes", "critics_score", "audience_score", "runtime"), 
                  selected = "critics_score"),
      
      # Select variable for color
      selectInput(inputId = "___", 
                  label = "____",
                  choices = c(___),
                  selected = "___")
    ),
    
    # Output: Show scatterplot
    mainPanel(
      plotOutput(outputId = "scatterplot")
    )
  )
)

# Define server function required to create the scatterplot
server <- function(input, output, session) {
  
  # Create the scatterplot object the plotOutput function is expecting
  output$scatterplot <- renderPlot({
    ggplot(data = movies, aes_string(x = input$x, y = input$y,
                                     color = ___)) +
      geom_point()
  })
}

# Create Shiny app object
shinyApp(ui = ui, server = server)
```

*** =solution
```{r}
library(shiny)
library(ggplot2)
load(url("http://s3.amazonaws.com/assets.datacamp.com/production/course_4850/datasets/movies.Rdata"))

# Define UI for application that plots features of movies
ui <- fluidPage(
  
  # Sidebar layout with a input and output definitions
  sidebarLayout(
    
    # Inputs: Select variables to plot
    sidebarPanel(
      
      # Select variable for y-axis
      selectInput(inputId = "y", 
                  label = "Y-axis:",
                  choices = c("imdb_rating", "imdb_num_votes", "critics_score", "audience_score", "runtime"), 
                  selected = "audience_score"),
      
      # Select variable for x-axis
      selectInput(inputId = "x", 
                  label = "X-axis:",
                  choices = c("imdb_rating", "imdb_num_votes", "critics_score", "audience_score", "runtime"), 
                  selected = "critics_score"),
      
      # Select variable for color
      selectInput(inputId = "z", 
                  label = "Color by:",
                  choices = c("title_type", "genre", "mpaa_rating", "critics_rating", "audience_rating"),
                  selected = "mpaa_rating")
    ),
    
    # Output: Show scatterplot
    mainPanel(
      plotOutput(outputId = "scatterplot")
    )
  )
)

# Define server function required to create the scatterplot
server <- function(input, output, session) {
  
  # Create the scatterplot object the plotOutput function is expecting
  output$scatterplot <- renderPlot({
    ggplot(data = movies, aes_string(x = input$x, y = input$y,
                                     color = input$z)) +
      geom_point()
  })
}

# Create a Shiny app object
shinyApp(ui = ui, server = server)
```

*** =sct
```{r}
# mention the inputs 
```

--- type:NormalExercise lang:r xp:100 skills:1 key:b1565759ca
## Extend the UI further

The potential variables the user can select for the `x` and `y` axes and `color` currently appear in the UI of the app the same way that they are spelled in the data frame header. However we might want to label them in a way that is more human readable. We can achieve this using named vectors for the `choices` argument, in the format of `"Human readable label" = "variable_name"`.

*** =instructions

- Fill in the blanks with human readable labels for `x` and `y` inputs.
- Re-create the `selectInput` widget for color, `z`, with options `"title_type"`, `"genre"`, `"mpaa_rating"`, `"critics_rating"`, and `"audience_rating"`, default selection `"mpaa_rating"` just like in the previous exercise, but this time use human readable labels as well.

*** =hint

*** =pre_exercise_code
```{r}
library(shiny)
library(ggplot2)
```

*** =sample_code
```{r}
library(shiny)
library(ggplot2)
load(url("http://s3.amazonaws.com/assets.datacamp.com/production/course_4850/datasets/movies.Rdata"))

# Define UI for application that plots features of movies
ui <- fluidPage(
  
  # Sidebar layout with a input and output definitions
  sidebarLayout(
    
    # Inputs: Select variables to plot
    sidebarPanel(
      
      # Select variable for y-axis
      selectInput(inputId = "y", 
                  label = "Y-axis:",
                  choices = c(___ = "imdb_rating", 
                              ___ = "imdb_num_votes", 
                              ___ = "critics_score", 
                              ___ = "audience_score", 
                              ___ = "runtime"), 
                  selected = "audience_score"),
      
      # Select variable for x-axis
      selectInput(inputId = "x", 
                  label = "X-axis:",
                  choices = c(___ = "imdb_rating", 
                              ___ = "imdb_num_votes", 
                              ___ = "critics_score", 
                              ___ = "audience_score", 
                              ___ = "runtime"), 
                  selected = "critics_score"),
      
      # Select variable for color
      selectInput(___)
    ),
    
    # Output: Show scatterplot
    mainPanel(
      plotOutput(outputId = "scatterplot")
    )
  )
)

# Define server function required to create the scatterplot
server <- function(input, output, session) {
  
  # Create the scatterplot object the plotOutput function is expecting
  output$scatterplot <- renderPlot({
    ggplot(data = movies, aes_string(x = input$x, y = input$y,
                                     color = input$z)) +
      geom_point()
  })
}

# Create a Shiny app object
shinyApp(ui = ui, server = server)
```

*** =solution
```{r}
library(shiny)
library(ggplot2)
load(url("http://s3.amazonaws.com/assets.datacamp.com/production/course_4850/datasets/movies.Rdata"))

# Define UI for application that plots features of movies
ui <- fluidPage(
  
  # Sidebar layout with a input and output definitions
  sidebarLayout(
    
    # Inputs: Select variables to plot
    sidebarPanel(
      
      # Select variable for y-axis
      selectInput(inputId = "y", 
                  label = "Y-axis:",
                  choices = c("IMDB rating"          = "imdb_rating", 
                              "IMDB number of votes" = "imdb_num_votes", 
                              "Critics score"        = "critics_score", 
                              "Audience score"       = "audience_score", 
                              "Runtime"              = "runtime"), 
                  selected = "audience_score"),
      
      # Select variable for x-axis
      selectInput(inputId = "x", 
                  label = "X-axis:",
                  choices = c("IMDB rating"          = "imdb_rating", 
                              "IMDB number of votes" = "imdb_num_votes", 
                              "Critics score"        = "critics_score", 
                              "Audience score"       = "audience_score", 
                              "Runtime"              = "runtime"), 
                  selected = "critics_score"),
      
      # Select variable for color
      selectInput(inputId = "z", 
                  label = "Color by:",
                  choices = c("Title type" = "title_type", 
                              "Genre" = "genre", 
                              "MPAA rating" = "mpaa_rating", 
                              "Critics rating" = "critics_rating", 
                              "Audience rating" = "audience_rating"),
                  selected = "mpaa_rating")
    ),
    
    # Output: Show scatterplot
    mainPanel(
      plotOutput(outputId = "scatterplot")
    )
  )
)

# Define server function required to create the scatterplot
server <- function(input, output, session) {
  
  # Create the scatterplot object the plotOutput function is expecting
  output$scatterplot <- renderPlot({
    ggplot(data = movies, aes_string(x = input$x, y = input$y,
                                     color = input$z)) +
      geom_point()
  })
}

# Create a Shiny app object
shinyApp(ui = ui, server = server)
```

*** =sct
```{r}
# mention the inputs 
```

--- type:VideoExercise lang:r xp:50 skills:1 key:fc9afde4cc
## Server function

01-03-server

*** =video_link
//player.vimeo.com/video/154783078

*** =video_hls
//videos.datacamp.com/transcoded/000_placeholders/v1/hls-temp.master.m3u8


--- type:MultipleChoiceExercise lang:r xp:50 skills:1 key:54f615ba72
## Rules of server functions

Which of the following is not true about server functions?

*** =instructions
- Server functions should include a call to `runApp()`
- Objects to be displayed should be saved to `output$`
- Reactive objects should be built with `render*()` functions
- Input values should be referred to with `input$`

*** =hint


*** =pre_exercise_code
```{r}
library(shiny)
```

*** =sct
```{r}
test_mc(1)
```


--- type:NormalExercise lang:r xp:100 skills:1 key:8197ac44a6
## Fix it up

On the right is the code for the Shiny app we built earlier, however currently the code is broken. Specifically there are errors in the definition of the server function as well as in the `mainPanel` of the UI.

*** =instructions

- Review the app and identify errors in the code
- Fix the errors and test out the app

*** =hint

*** =pre_exercise_code
```{r}
library(shiny)
library(ggplot2)
```

*** =sample_code
```{r}
library(shiny)
library(ggplot2)
load(url("http://s3.amazonaws.com/assets.datacamp.com/production/course_4850/datasets/movies.Rdata"))

# Define UI for application that plots features of movies
ui <- fluidPage(
  
  # Sidebar layout with a input and output definitions
  sidebarLayout(
    
    # Inputs: Select variables to plot
    sidebarPanel(
      
      # Select variable for y-axis
      selectInput(inputId = "y", 
                  label = "Y-axis:",
                  choices = c("IMDB rating"          = "imdb_rating", 
                              "IMDB number of votes" = "imdb_num_votes", 
                              "Critics score"        = "critics_score", 
                              "Audience score"       = "audience_score", 
                              "Runtime"              = "runtime"), 
                  selected = "audience_score"),
      
      # Select variable for x-axis
      selectInput(inputId = "x", 
                  label = "X-axis:",
                  choices = c("IMDB rating"          = "imdb_rating", 
                              "IMDB number of votes" = "imdb_num_votes", 
                              "Critics score"        = "critics_score", 
                              "Audience score"       = "audience_score", 
                              "Runtime"              = "runtime"), 
                  selected = "critics_score"),
      
      # Select variable for color
      selectInput(inputId = "z", 
                  label = "Color by:",
                  choices = c("Title type" = "title_type", 
                              "Genre" = "genre", 
                              "MPAA rating" = "mpaa_rating", 
                              "Critics rating" = "critics_rating", 
                              "Audience rating" = "audience_rating"),
                  selected = "mpaa_rating")
    ),
    
    # Output: Show scatterplot
    mainPanel(
      plotOutput(outputId = "scatterPlot")
    )
  )
)

# Define server function required to create the scatterplot
server <- function(input, output, session) {
  
  # Create the scatterplot object the plotOutput function is expecting
  output$scatterplot <- renderTable({
    ggplot(data = movies, aes_string(x = x, y = y,
                                     color = z)) +
      geom_point()
  })
}

# Create a Shiny app object
shinyApp(ui = ui, server = server)
```

*** =solution
```{r}
library(shiny)
library(ggplot2)
load(url("http://s3.amazonaws.com/assets.datacamp.com/production/course_4850/datasets/movies.Rdata"))

# Define UI for application that plots features of movies
ui <- fluidPage(
  
  # Sidebar layout with a input and output definitions
  sidebarLayout(
    
    # Inputs: Select variables to plot
    sidebarPanel(
      
      # Select variable for y-axis
      selectInput(inputId = "y", 
                  label = "Y-axis:",
                  choices = c("IMDB rating"          = "imdb_rating", 
                              "IMDB number of votes" = "imdb_num_votes", 
                              "Critics score"        = "critics_score", 
                              "Audience score"       = "audience_score", 
                              "Runtime"              = "runtime"), 
                  selected = "audience_score"),
      
      # Select variable for x-axis
      selectInput(inputId = "x", 
                  label = "X-axis:",
                  choices = c("IMDB rating"          = "imdb_rating", 
                              "IMDB number of votes" = "imdb_num_votes", 
                              "Critics score"        = "critics_score", 
                              "Audience score"       = "audience_score", 
                              "Runtime"              = "runtime"), 
                  selected = "critics_score"),
      
      # Select variable for color
      selectInput(inputId = "z", 
                  label = "Color by:",
                  choices = c("Title type" = "title_type", 
                              "Genre" = "genre", 
                              "MPAA rating" = "mpaa_rating", 
                              "Critics rating" = "critics_rating", 
                              "Audience rating" = "audience_rating"),
                  selected = "mpaa_rating")
    ),
    
    # Output: Show scatterplot
    mainPanel(
      plotOutput(outputId = "scatterplot")
    )
  )
)

# Define server function required to create the scatterplot
server <- function(input, output, session) {
  
  # Create the scatterplot object the plotOutput function is expecting
  output$scatterplot <- renderPlot({
    ggplot(data = movies, aes_string(x = input$x, y = input$y,
                                     color = input$z)) +
      geom_point()
  })
}

# Create a Shiny app object
shinyApp(ui = ui, server = server)
```

*** =sct
```{r}

```
--- type:VideoExercise lang:r xp:50 skills:1 key:723aa768a4
## Recap

01-04-recap

*** =video_link
//player.vimeo.com/video/154783078

*** =video_hls
//videos.datacamp.com/transcoded/000_placeholders/v1/hls-temp.master.m3u8




